\documentclass{article}
\usepackage[utf8]{inputenc}

\begin{document}

\begin{titlepage}
\centering
\vspace*{6cm}
{\Large\bfseries TW Assignment \par}
\vspace{2cm}
{\Large Name: \textbf{Muhammad Saad} \par}
{\Large Registration Number: \textbf{21pwcse1997} \par}
{\Large Section: \textbf{B} \par}
\vspace{2cm}
{\Large Date: \textbf{\today} \par}
\end{titlepage}

\newpage

\section{INVESTIGATION OF SERVERLESS ARCHITECTURE}
\subsection{Summary}
This research paper is about exploring the serverless architecture. Serverless is a cloud computing model where clients don't have to write code for a server but use cloud APIs that does everything. All the data manipulations like create, read, update and delete, is handled by the cloud. It is also possible to run custom code by the client on the cloud.

\subsection{Main Contributions}
\begin{itemize}
  \item  Pay-as-you-go model reduces idle costs and initial investment.
  \item Requires additional knowledge and skills for effective use.
  \item May not perform well for long, compute-intensive tasks.
	\item  Differences in services across cloud providers can complicate transitions.
\end{itemize}

\subsection{Limitations}
\begin{itemize}
  \item Dependency on a specific cloud provider can limit flexibility and control.
  \item Decomposition into multiple independent parts can increase the attack surface.
  \item Requires additional knowledge and skills for effective use.
	\item May not perform well for long, compute-intensive tasks.
	\item Differences in services across cloud providers can complicate transitions.
\end{itemize}

\end{document}
